Sound Automated Synthesis of Digital Controllers

### 2016-10-09 paper/paper.tex
created new section on fixed point manipulation

@@

Let $\tilde C(z)$ and $\tilde G(z)$ be the transfer functions represented
using a Fixed-Point Finite Word Length with integer size $I$ and fractional
size $F$.  We shall denote this domain as $\mathbb{R}\langle I,F \rangle$ and obtain 
%
\begin{align}
\small
\label{digital_controller_plant_tf}
\tilde C(z)&=\frac{\tilde \beta_{0}+\tilde \beta_{1}z^{-1}+...+\tilde \beta_{M_C}z^{-M_C}}{\tilde \alpha_{0}+\tilde \alpha_{1}z^{-1}+...+\tilde \alpha_{N_C}z^{-N_C}}, \\
%\label{plant_tf}
\tilde G(z)&=\frac{\tilde b_{0}+\tilde b_{1}z^{-1}+...+\tilde b_{M_G}z^{-M_G}}{\tilde a_{0}+\tilde a_{1}z^{-1}+...+\tilde a_{N_G}z^{-N_G}}.
\end{align}

\textcolor{blue}{The digital controller $C(z)$ also suffers disturbances from the FWL effects, with roundoffs in coefficients that may change the closed-loop poles and zeros position, and consequently affect its stability, as shown in~\cite{Bessa16}.}

\textcolor{blue}{Since the controller is implemented in the $\mathbb{R}\langle I,F \rangle$ 
domain, {\it i.e.} a Fixed-Point Finite Word Length with integer size $I$ and fractional
size $F$, we shall denote this domain as $\mathbb{R}\langle I,F \rangle$ and assume that there exists a function $FWL[\cdot]:\mathbb{R}^{n}\rightarrow \mathbb{R}^{n}\langle I,F \rangle$, which applies the FWL effects to a polynomial. Thus the perturbed controller model $\hat{C}(z)$ may be computed from the original model $C(z)=\frac{N_{C}(z)}{D_{C}(z)}$ using the following expression:
\begin{equation}
\hat{C}(z)=\frac{FWL[N_{C}(z)]}{FWL[D_{C}(z)]}
\end{equation}}

@@

Since the controller is implemented in the $\mathbb{R}\langle I,F \rangle$
domain,
%
 $$\forall i \leq N_C, j \leq M_C\ \  \tilde \beta_{i} \equiv \beta_{i} \wedge \tilde \alpha_{j} \equiv \alpha_{j} \Leftrightarrow \tilde C(z) \equiv C(z)$$
%
However, given a real plant and real controller $\hat{G}(z), \hat{C}(z)$, and a discretizing function
\begin{align}
\small
\label{FWL_tf}
&f (x \in \mathbb{R}) \rightarrow \tilde x \in \mathbb{R}\langle I,F \rangle \triangleq \tilde x = x-(x\mod\tilde x)
\end{align}
we calculate $\tilde C(z)=f(\hat{C}(z))$ and $\tilde G(z)=f(\hat{G}(z))$
\begin{align}
\label{digital_controller_tf}
\tilde C(z)&=\frac{(\hat{\beta}_{0}+\Delta_q \hat{\beta}_{0}) +...+(\hat{\beta}_{M_C}+\Delta_q \hat{\beta}_{M_C})z^{-M_C}}{(\hat{\alpha}_{0}+\Delta_q \hat{\alpha}_{0})+...+(\hat{\alpha}_{N_C}+\Delta_q \hat{\alpha}_{N_C})z^{-N_G}} \nonumber \\
&= C(z) =\hat{C}(z) + \Delta{C}(z)\\
\label{digital_plant_tf}
\tilde G(z)&=\frac{(\hat{b}_{0}+\Delta_q \hat{b}_{0}) +...+(\hat{b}_{M_G}+\Delta_q \hat{b}_{M_G})z^{-M_G}}{(\hat{a}_{0}+\Delta_q \hat{a}_{0})+...+(\hat{a}_{N_G}+\Delta_q \hat{a}_{N_G})z^{-N_G}} \nonumber \\
&=G(z)+\Delta{G}_P(z)+\Delta{G}_Q(z).
\end{align}
where $\Delta_q\hat{c}_i=\hat{c}_i\mod \tilde{c}_i$
\red{[please, how can you claim that the error enters as written? Can you show the derivation? Can you show a reference that has derived this form?]}
\green{[the definition of f (now included) should explain it clearly]}
and $\Delta{G}_Q(z)$ represents the plant uncertainty caused by
the rounding off effect.  We may encompass all the uncertainty as
$\Delta{G}(z)=\Delta{G}_P(z)+\Delta{G}_Q(z)$.  Since this value is
nondeterministic, the set of possible polynomial coefficients of $\tilde G(z)$ is
denoted by $\mathfrak{G}(z)$.

%------------------------------------
\subsection{DSVerifier Verification Flow}
\label{verification-flow}
%------------------------------------

>>>>

\textcolor{green}{DC: I believe Alessandro wanted to be more specific about the nature of the modelling error which is why I was trying to explain it. In this case I believe there are also two separate sources for $\Delta{G}_p(z)$. The first is $\Delta{G}_p(s)$ (which is basically a known tolerance that may come from G(t)), whilst the second is the error introduced by the calculations in the $\mathcal{Z}$ transform. Alessandro, do you want to address this second term (possibly as explained below) or just leave the blue section as is?}

@@

\textcolor{blue}{Since the controller is implemented using a FWL, $C(z)$ also suffers disturbances from the FWL effects, with roundoffs in coefficients that may change the closed-loop poles and zeros position, and consequently affect its stability, as shown in~\cite{Bessa16}.}

\textcolor{blue}{
Let $\hat{C}(z)$ be the transfer function of the digital controller represented using this FWL with integer size $I$ and fractional
size $F$.  We shall denote this domain as $\mathbb{R}\langle I,F \rangle$
%Since the controller is implemented in the $\mathbb{R}\langle I,F \rangle$ 
%domain, {\it i.e.} a Fixed-Point Finite Word Length with integer size $I$ and fractional
%size $F$, we shall denote this domain as $\mathbb{R}\langle I,F \rangle$ 
and assume that there exists a function $FWL[\cdot]:\mathbb{R}^{n}\rightarrow \mathbb{R}^{n}\langle I,F \rangle$, which applies the FWL effects to a polynomial. Thus the perturbed controller model $\hat{C}(z)$ may be computed from the original model $C(z)=\frac{N_{C}(z)}{D_{C}(z)}$ using the following expression:
\begin{equation}
\hat{C}(z)=\frac{FWL[N_{C}(z)]}{FWL[D_{C}(z)]}
\end{equation}}

@@

\textcolor{green}{[DC: I changed it to a separate section to differentiate the effects, so we should keep both]}

%------------------------------------
\subsection{Using Fixed Point Arithmetic for verification}
\label{fixedpoint_verification}
%------------------------------------

Let $\tilde C(z)$ and $\tilde G(z)$ be the transfer functions represented
using Fixed-Point bit vectors with integer size $I_b$ and fractional
size $F_b$, whose domain is $\mathbb{R}\langle I_b,F_b \rangle \supseteq \mathbb{R}\langle I,F \rangle$.
%
\begin{align}
\small
\label{digital_controller_plant_tf}
\tilde C(z)&=\frac{\tilde \beta_{0}+\tilde \beta_{1}z^{-1}+...+\tilde \beta_{M_C}z^{-M_C}}{\tilde \alpha_{0}+\tilde \alpha_{1}z^{-1}+...+\tilde \alpha_{N_C}z^{-N_C}}, \\
%\label{plant_tf}
\tilde G(z)&=\frac{\tilde b_{0}+\tilde b_{1}z^{-1}+...+\tilde b_{M_G}z^{-M_G}}{\tilde a_{0}+\tilde a_{1}z^{-1}+...+\tilde a_{N_G}z^{-N_G}}.
\end{align}

Since the controller is synthesised in the $\mathbb{R}\langle I,F \rangle$
domain,
%
 $$\forall i \leq N_C, j \leq M_C\ \  \tilde \beta_{i} \equiv \beta_{i} \wedge \tilde \alpha_{j} \equiv \alpha_{j} \Leftrightarrow \tilde C(z) \equiv C(z)$$
%
However, given a real plant and real controller $\hat{G}(z), \hat{C}(z)$, and a discretizing function
\begin{align}
\small
\label{FWL_tf}
&f_b (x \in \mathbb{R}) \rightarrow \tilde x \in \mathbb{R}\langle I_b,F_b \rangle \triangleq \tilde x = x-(x\mod\tilde x)
\end{align}
we calculate $\tilde C(z)=f(\hat{C}(z))$ and $\tilde G(z)=f(\hat{G}(z))$
\begin{align}
\label{digital_controller_tf}
\tilde C(z)&=\frac{(\hat{\beta}_{0}+\Delta_q \hat{\beta}_{0}) +...+(\hat{\beta}_{M_C}+\Delta_q \hat{\beta}_{M_C})z^{-M_C}}{(\hat{\alpha}_{0}+\Delta_q \hat{\alpha}_{0})+...+(\hat{\alpha}_{N_C}+\Delta_q \hat{\alpha}_{N_C})z^{-N_G}} \nonumber \\
&= C(z) =\hat{C}(z) + \Delta{C}_b(z)\\
\label{digital_plant_tf}
\tilde G(z)&=\frac{(\hat{b}_{0}+\Delta_q \hat{b}_{0}) +...+(\hat{b}_{M_G}+\Delta_q \hat{b}_{M_G})z^{-M_G}}{(\hat{a}_{0}+\Delta_q \hat{a}_{0})+...+(\hat{a}_{N_G}+\Delta_q \hat{a}_{N_G})z^{-N_G}} \nonumber \\
&=G(z)+\Delta{G}_P(z)+\Delta{G}_b(z).
\end{align}
where $\Delta_b\hat{c}_i=\hat{c}_i\mod \tilde{c}_i$
and $\Delta{G}_b(z)$ represents the plant uncertainty caused by
the rounding off effect.  We may encompass all the uncertainty as
$\Delta{G}(z)=\Delta{G}_P(z)+\Delta{G}_b(z)$.  Since this value is
nondeterministic, the set of possible polynomial coefficients of $\tilde G(z)$ is
denoted by $\mathfrak{G}(z)$.

Two approaches have been taken to synthesise a stabe controller based on this model.
The first uses a nondeterministic plant selected from $\mathfrak{G}(z)$ where the expected error of the bit vector calculations is upper bound at the source by selecting a conservative $\Delta{G}_b(z)$ that encompasses its propagation over a number of operations. The verification phase produces a counterexample if some instance of $\mathfrak{G}(z)$ does not verify the properties and the CEGIS loop iteratively looks for a new controller that verifies the supplied counterexamples.
Although it is fast, the conservative error calculation can be restrictive to a small number of benchmarks.

The second approch uses interval arithmetic, representing the coefficients $[{c}_i-\Delta_p{c}_i-\Delta_b{c}_i\ ,\ {c}_i+\Delta_p{c}_i-\Delta_b{c}_i+(2^{-F_b})]$ 
and rounding operations outwards during the process. This results in a tighter calculation of the error at the cost of performance.
%------------------------------------
\subsection{DSVerifier Verification Flow}
\label{verification-flow}
%------------------------------------
